---
title: "AutoEpi Daily Syndrome Report"
subtitle: "Automated RMarkdown Template"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: false
    theme: readable
    code_folding: hide
    df_print: paged
params:
  date_key: "20301"
  syndrome_key: null
  report: !r list()
  meta: !r list()
---

<!--
This template expects to be rendered by a driver script that passes:
- params$report          (list: $count, $per10k, $maps each containing htmlwidgets or NULL)
- params$meta            (list: climate_table, climate_stats, aqi_stations, aqi_summary, aqi_plot)
Optionally:
- params$date_key        (character; key from autoepi_report list)
- params$syndrome_key    (character; name/key used in autoepi_report list)
The template will try to infer date and syndrome from whatever is available.
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, fig.align = "center")
`%||%` <- function(a, b) if (!is.null(a) && length(a)) a else b
has <- function(x) !is.null(x) && length(x) > 0
safekable <- function(df, caption = NULL) {
  if (!is.null(df) && NROW(df) > 0) {
    knitr::kable(df, caption = caption, format = "html")
  }
}

# Try to extract a human-readable title from a plotly htmlwidget
pluck_plot_title <- function(w) {
  tryCatch({
    if (inherits(w, "htmlwidget")) {
      # plotly stores layout$title$text in x$layout$title$text
      t1 <- w[["x"]][["layout"]][["title"]]
      if (is.character(t1)) return(t1)
      if (is.list(t1) && !is.null(t1$text)) return(as.character(t1$text))
    }
    NULL
  }, error = function(e) NULL)
}

# Parse a date-like thing to "MM/DD/YYYY" format
fmt_date_label <- function(d) {
  if (inherits(d, "Date")) return(format(d, "%m/%d/%Y"))
  if (is.numeric(d) && !is.na(d)) {
    s <- as.character(d)
    if (nchar(s) == 8) return(format(as.Date(s, "%Y%m%d"), "%m/%d/%Y"))
  }
  if (is.character(d) && nzchar(d)) {
    s <- d[1]
    # Accept common forms and convert to MM/DD/YYYY
    if (grepl("^\\d{4}-\\d{2}-\\d{2}$", s)) return(format(as.Date(s, "%Y-%m-%d"), "%m/%d/%Y"))
    if (grepl("^\\d{8}$", s)) return(format(as.Date(s, "%Y%m%d"), "%m/%d/%Y"))
    if (grepl("^\\d{2}/\\d{2}/\\d{4}$", s)) {
      # mm/dd/yyyy -> keep as is
      return(s)
    }
  }
  NA_character_
}

# Heuristic date inference:
# 1) meta$climate_stats$Date
# 2) meta$aqi_summary$Date
# 3) first(meta$aqi_stations$Date)
# 4) fall back to params$date_key (string normalization attempted)
infer_date_label <- function(meta, date_key) {
  d <- tryCatch(meta$climate_stats$Date, error = function(e) NULL)
  lbl <- fmt_date_label(d)
  if (!is.na(lbl)) return(lbl)
  d <- tryCatch(meta$aqi_summary$Date, error = function(e) NULL)
  lbl <- fmt_date_label(d)
  if (!is.na(lbl)) return(lbl)
  d <- tryCatch(meta$aqi_stations$Date, error = function(e) NULL)
  if (has(d)) {
    lbl <- fmt_date_label(d[1])
    if (!is.na(lbl)) return(lbl)
  }
  # Fall back to date_key if provided
  fmt_date_label(date_key) %||% as.character(date_key %||% "Unknown")
}

# Heuristic syndrome inference:
# Prefer explicit param; otherwise sniff plot titles for a leading name.
infer_syndrome <- function(report, syn_key) {
  if (has(syn_key)) return(as.character(syn_key))
  # Try count widgets first
  cw <- tryCatch(report$count, error = function(e) NULL)
  widgets <- c(
    cw$age, cw$age_gender, cw$sex, cw$race, cw$ethnicity,
    tryCatch(report$per10k$age, error = function(e) NULL),
    tryCatch(report$per10k$sex, error = function(e) NULL),
    tryCatch(report$per10k$hospital, error = function(e) NULL)
  )
  titles <- vapply(widgets, pluck_plot_title, FUN.VALUE = character(1))
  titles <- titles[nzchar(titles)]
  if (length(titles)) {
    # Often titles are like "Age Group Distribution - <SYNDROME> (<DATE>)"
    # Pull the segment between ' - ' and ' (' if present
    s <- titles[1]
    m <- regmatches(s, regexec(" -\\s*([^\\(]+)\\s*\\(", s))[[1]]
    if (length(m) >= 2) return(trimws(m[2]))
    # Else try end after last ' - '
    parts <- unlist(strsplit(s, " - "))
    if (length(parts) >= 2) return(trimws(parts[length(parts)]))
    return(s) # fallback raw title
  }
  "Unknown Syndrome"
}

# Climate station name mapping function
map_station_name <- function(station_raw) {
  station_map <- c(
    "pdt-eln" = "Ellensburg, Washington",
    "otx-lws" = "Lewiston, Idaho",
    "pdt-dls" = "Dallesport, Washington",
    "pdt-psc" = "Pasco, Washington",
    "sew-sew" = "Seattle, Washington",
    "otx-geg" = "Spokane, Washington",
    "pqr-vuo" = "Vancouver, Washington",
    "pdt-alw" = "Walla Walla, Washington",
    "otx-eat" = "Wenatchee, Washington",
    "pdt-ykm" = "Yakima, Washington",
    "sew-olm" = "Olympia, Washington",
    "otx-omk" = "Omak, Washington",
    "sew-bli" = "Bellingham, Washington",
    "otx-dew" = "Deer Park, Washington"
  )
  # Convert input to lowercase for case-insensitive matching
  station_raw_lower <- tolower(station_raw)
  # Find the matching key (also converted to lowercase for comparison)
  matched_key <- names(station_map)[tolower(names(station_map)) == station_raw_lower]
  ifelse(length(matched_key) > 0, station_map[matched_key[1]], station_raw)
}

# AQI station name mapping function
map_aqi_station_name <- function(station_id) {
  # AQI stations typically have numeric IDs or different naming conventions
  # For now, return a more descriptive name based on the ID
  # This can be expanded with actual AQI station mappings when available
  if (is.null(station_id) || is.na(station_id)) {
    return("Unknown Station")
  }
  
  # If it's already a descriptive name, return as-is
  if (grepl("[A-Za-z]", station_id)) {
    return(as.character(station_id))
  }
  
  # For numeric IDs, create a descriptive name
  paste("AQI Station", as.character(station_id))
}

# Age group ordering function
order_age_groups <- function(age_groups) {
  # Define the correct order for age groups
  age_order <- c("0-5", "6-18", "19-24", "25-34", "35-44", "45-54", "55-64", "65+", "Unknown")
  # Filter to only include age groups that exist in the data
  existing_ages <- intersect(age_order, age_groups)
  # Add any remaining age groups that weren't in our predefined order
  remaining_ages <- setdiff(age_groups, age_order)
  c(existing_ages, remaining_ages)
}

# Materialize our inferred labels
.date_label <- infer_date_label(params$meta %||% list(), params$date_key %||% NA_character_)
.syndrome_label <- infer_syndrome(params$report %||% list(), params$syndrome_key %||% NULL)
```

<style>
/* Minimal, readable table and section spacing tweaks */
table { font-size: 1.1rem; width: 100%; }
h1, h2, h3 { margin-top: 1.4em; }
.figure, .html-widget { margin-top: 0.6em; margin-bottom: 1.1em; }
.smallcaps { font-variant: small-caps; color: #555; }
.badge { display: inline-block; padding: 0.2em 0.5em; background: #efefef; border-radius: 4px; margin-right: 0.3em; }
.kable-table { width: 100% !important; }
</style>

# Report Overview

- **Date (inferred):** `r .date_label`
- **Date (key):** `r params$date_key %||% "N/A"`  
- **Syndrome (inferred):** `r .syndrome_label`
- **Syndrome (key):** `r params$syndrome_key %||% "N/A"`  
- **Generated:** `r format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z")`

This report compiles counts, per-10k rates, and an optional ZIP/ZCTA choropleth for the selected syndrome and date. It also includes per-date meta for climate and air quality (PM2.5 24-hr).

---

# Climate 

```{r climate, results="asis"}
clim_tab   <- params$meta$climate_table %||% NULL
clim_stats <- params$meta$climate_stats %||% NULL

if (has(clim_stats)) {
  cat("## Climate Summary\n")
  print(safekable(clim_stats, caption = "Climate Averages (across included stations)"))
}

if (has(clim_tab)) {
  cat("## Climate Stations\n")
  # Debug: print what station IDs we're working with
  if (!is.null(clim_tab) && "StationRaw" %in% names(clim_tab)) {
    message("Debug - Climate station IDs: ", paste(unique(clim_tab$StationRaw), collapse=", "))
    
    # Keep the station IDs but rename the column for display
    clim_tab_display <- clim_tab %>% 
      rename(`Climate Station` = StationRaw)
  } else {
    clim_tab_display <- clim_tab
  }
  print(safekable(clim_tab_display, caption = "Station Readings"))
}
```

# Air Quality

```{r aqi, results="asis"}
aqi_plot    <- params$meta$aqi_plot %||% NULL
aqi_summary <- params$meta$aqi_summary %||% NULL
aqi_stn     <- params$meta$aqi_stations %||% NULL

if (has(aqi_plot)) {
  cat("## AQI PM2.5 (24-hr)\n")
  aqi_plot
} else if (has(aqi_summary)) {
  cat("## AQI PM2.5 (24-hr) Summary\n")
  print(safekable(aqi_summary, caption = "County-Level Summary"))
}

if (has(aqi_stn)) {
  cat("### AQI Stations\n")
  # Debug: print what columns are available in AQI data
  message("Debug - AQI columns: ", paste(names(aqi_stn), collapse=", "))
  if ("SiteName" %in% names(aqi_stn)) {
    message("Debug - AQI SiteName values: ", paste(unique(aqi_stn$SiteName), collapse=", "))
  }
  if ("StationID" %in% names(aqi_stn)) {
    message("Debug - AQI StationID values: ", paste(unique(aqi_stn$StationID), collapse=", "))
  }
  
  # Remove station name columns and AQSID, keep original data
  if (!is.null(aqi_stn)) {
    aqi_stn_display <- aqi_stn %>%
      select(-any_of(c("StationName", "SiteName", "AQSID")))
  } else {
    aqi_stn_display <- aqi_stn
  }
  print(safekable(aqi_stn_display, caption = "Station Measurements"))
}
```

# Counts

The following sections display counts by demographic or facility dimension. If an element is missing in the input object, it will be omitted automatically.

```{r counts, results="asis"}
cnt <- params$report$count %||% list()

if (has(cnt$age)) {
  cat("## Age Group (Count)\n")
  cnt$age
}
if (has(cnt$age_gender)) {
  cat("## Age × Gender (Count)\n")
  cnt$age_gender
}
if (has(cnt$sex)) {
  cat("## Sex (Count)\n")
  cnt$sex
}
if (has(cnt$race)) {
  cat("## Race (Count)\n")
  cnt$race
}
if (has(cnt$ethnicity)) {
  cat("## Ethnicity (Count)\n")
  cnt$ethnicity
}
```

# Per-10k Rates

Rates are computed upstream using the all-visits denominators for the report date. Sections render only if corresponding widgets exist.

```{r per10k, results="asis"}
r10 <- params$report$per10k %||% list()

if (has(r10$age)) {
  cat("## Age Group (per 10k)\n")
  r10$age
}
if (has(r10$sex)) {
  cat("## Sex (per 10k)\n")
  r10$sex
}
if (has(r10$hospital)) {
  cat("## Hospital (per 10k)\n")
  r10$hospital
}
```

# Map

If available, the map displays counts by ZIP/ZCTA for the report date and syndrome. Polygons with zero counts are rendered with low opacity to retain geographic context.

```{r maps, results="asis"}
mp <- params$report$maps %||% list()
if (has(mp$choropleth)) {
  cat("## ZIP/ZCTA Choropleth\n")
  mp$choropleth
}
```

# Notes & Methods

- Data source: NSSP ESSENCE (as configured in the generating pipeline).
- Date scope: Single date per report; meta (climate/AQI) is per-date and reused for all syndromes on that date.
- Demographics: Age groups are derived from settings. Unknown/empty categories may be omitted upstream.
- Per-10k rates: Numerators from the selected syndrome subset; denominators from all visits on the same date.
- Mapping: ZIP to ZCTA crosswalk (if configured) is applied upstream to drop PO Box/unique ZIPs and aggregate to ZCTAs before joining to geometry.
- Reproducibility: All visualizations are pre-built htmlwidgets; this template does not re-query external services.

## Quality Flags (optional)

If you pass any flags via the driver, consider printing them here for auditor notes (e.g., missing denominators, dropped ZIPs, schema drift).

# Appendix

```{r session, echo=FALSE}
sessionInfo()
```
